Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-22 drafts/hybrid_zone.tex
minor fixes to intro

\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}. While these patterns of co-ancestry have been described in multi-locus systems, the  \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \cite[e.g.][]{Gravel2012,Sedghifar2015}.  Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.

Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \cite{Barton1979}. 
%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim here  is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}}.

@@

Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions. We therefore expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.  This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.

Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}

>>>>

\yb{To the extent that  genetic variation can be exchanged between species via hybridization, both the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} Subsets of the genome that are under selection, however, can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986,Barton1983}. While these patterns of co-ancestry have been described in multi-locus systems, the  \yb{expected length of ancestral tracts around selected loci in admixing populations remains poorly understood}\alisa{, especially in the geographic context of hybrid zones.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place between every generation, forming chromosomal \emph{junctions} in ancestry \citep{Fisher1954, Chapman2002, Baird2003}.   
The expected distribution of lengths of unbroken ancestry in admixed populations has been previously described under neutral conditions both in single populations and in contact zones \citep[e.g.][]{Gravel2012,Sedghifar2015}.  Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection.

Variation in \yb{ the action of } selection against migrant alleles within the genome results in genome-wide heterogeneity in cline width, as the frequency of foreign ancestry surrounding targets of selection is reduced \citep{Barton1979}. 
%The effect on patterns of ancestral block lengths is, however, less well explored theoretically and \yb{empirically.
\yb{While empirical patterns of the block length distribution in hybrid zones have rarely been used to inform the history of selection,}  \alisa{the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \citep{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim here  is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}}.

@@

Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions. We therefore expect \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. As a corollary to this, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.  This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.

Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
